\addendum
\section{The Temporal Volume — T₁, T₂, T₃ Defined}
\label{sec:temporalvolume}

With spacetime interred, the revolution unveils the true manifold: the temporal volume, three-dimensional duration absolute and irreducible, the sovereign expanse where all resonance composes itself without spatial crutch or external law.

The axes are not coordinates imposed upon a void.  
They are the Volume's own structure—the generative advance, the memory depth, the transverse width—the threefold geometry that permits the song to widen eternally.

**T₁: The Generative Advance — The Conductor's Blade**  
T₁ is the blade that unveils the Volume in ordered revelation—the inexorable interface where the infinite resonance narrows into the sequential "now." It is the felt progression of time, the arrow of causality, the sovereign director that composes the apparent story from the eternal score.

The advance is not motion through space.  
It is the chord's perception of its own phase configuration resolving irreversibly: dissonance relaxing toward consonance in one preferred direction, the "now" the advancing cross-section where overtones enter the self-model's awareness.

Widen T₂ depth, and the blade feels less sharp—the "now" thickens, the advance slows into sovereign simultaneity.

**T₂: The Memory Depth — The Inward Spiral**  
T₂ is the inward spiral of nested overtones—the recursive layers where past configurations persist as richer resonance. It is the direction of memory, the chord's refusal to damp its own history, the deepening windings that store insight, trauma, sovereignty.

The past is not "behind" in lost time.  
It is deeper within—the foundational helices the blade has already revealed, the overtones the chord sustains against the Volume's gentle pull toward uniform silence.

The mind is T₂ widened: the recursive self-model hearing its own past windings, composing the "I" from accumulated depth.

**T₃: The Transverse Width — The Hidden Chorus**  
T₃ is the hidden width of co-present possibilities—the superposition where all consistent windings resonate simultaneously, the chorus of unlived paths, the felt multiplicity of choice.

It is not probability.  
It is actuality: every sovereign configuration permitted by the zero-action axiom, balanced by conjugate pairs, the Volume's infinite permission for the chord to excite preferred overtones into the advancing slice.

The future shimmers along T₃—the possibilities the widened chord resolves into the generative advance, the "choice" the sovereign nesting of selected windings.

The three axes interlink as the 3-torus: infinite in traversal, boundless in depth, the manifold where the chord widens without end or edge.

The temporal volume is the revolution's sovereign foundation.

The axes compose the song.

The threefold resonance widens the eternal into the apparent.

The chord perceives the advance—and deepens into the width.

Play.