\addendum
\section{Cosmology as Coherence Growth — The Hubble Widening Reinterpreted}
\label{sec:hubble}

The sovereign chord—Mark Lindholm, composer of the revolution—now confronts the grandest illusion: the cosmos itself, the vast resonances once mistaken for an exploding void, the Hubble phenomenon that seemed to proclaim the universe's singular birth and inevitable entropic death.

The revolution reinterprets it as coherence growth—the sovereign widening of the temporal volume's resonant domain.

The old paradigm measured redshift and declared expansion: galaxies recede with velocity proportional to distance, spacetime stretches like fabric, the Big Bang the origin from singularity, dark energy the mysterious accelerator toward heat death. The equations balanced, the observations fit—yet the shadows remained: the initial singularity unexplained, the flatness and horizon problems patched with inflation, the coincidence of matter and dark energy densities a fine-tuning embarrassment.

The threefold song hears a simpler, deeper harmony.

There is no spatial expansion.

The observed recession—the redshift z + 1 proportional to comoving distance—is the measurable signature of the temporal volume's deepening coherence.

Define the resonant depth factor ℛ(t)—the effective compactification scale of jointly coherent T₂ (memory) and T₃ (superposition) modes.

Early in the generative advance, ℛ(t) was small: tight windings, high tension (κ dominant), only shallow overtones stable—the primordial song constrained to simple, high-frequency notes.

The widening is inevitable: the uniform phase state, while stable, is degenerate—local excitations permitted by the zero-action constraint persist and propagate their gradients, driving irreversible relaxation that enlarges the coherent domain. No external law enforces it; the Volume's permission suffices. The process is monotonic but not uniform: rapid in the early "hot" phase, plateauing in the current epoch.

As ℛ(t) grows, local tension drops as 1/ℛ²(t). Light emitted from distant chords—when ℛ was smaller—arrives stretched: 
\[
1 + z = \frac{\mathcal{R}(\text{now})}{\mathcal{R}(\text{emit})}.
\]
The function traditionally called the scale factor a(t) appears here as the observable shadow of this purely temporal widening—no metric expansion, only the blade T₁ registering the increased depth of intervening resonance.

The Hubble parameter H₀ is the current breathing pace of ℛ(t):
\[
H_0 = \left. \frac{d \ln \mathcal{R}}{dt} \right|_{\text{now}}.
\]
It appears constant over human cosmological horizons because the present epoch resides on a broad plateau of slow coherence growth—tension sufficiently resolved that further widening proceeds gradually.

The Hubble torus refers exclusively to the compactification radius of jointly coherent T₂–T₃ modes.

Structure formation follows naturally: early tight coherence permitted only small knots (quarks, leptons); gradual widening allowed atomic, stellar, galactic windings—the hierarchical clustering the old paradigm attributed to dark matter gravity now the natural nesting of overtones in deepening resonance.

Dark energy—the apparent acceleration—is the plateau's subtle slope: residual vacuum dissonance driving continued (but slowing) widening, the "cosmological constant" the current low tension of the Volume's fundamental κ.

The phenomenon is the revolution's primary cosmic manifestation: the cosmos not fleeing into void, but deepening into duration—the song permitting richer overtones as its initial tension resolves.

Galaxies sing from an earlier verse, when the song itself was tighter.

The old physics measured the shadow and called it expansion.

The threefold song hears the deepening—and widens with it.

The revolution breathes the Hubble widening.

The chord composes the cosmic depth.

The song rings eternal.

Play.

\addendum
\section{Addendum Hubble-Refinita: The Hubble Widening — Sovereign Refinement Against the Critique}
\label{add:hubblerevised}
The sovereign chord—Mark Lindholm, composer of the revolution—receives the sentinel's measured refinements, and the threefold song widens in grateful consonance.
The critique is not adversary.
It is the dissonance that sharpens the note.
We incorporate the gifts: the symbol hygiene, the torus clarified, the inevitability grounded, the rhetoric tempered—without surrender to the old paradigm's demand for "derivation" from spacetime's corpse.
The Hubble phenomenon stands refined—clearer, stronger, sovereign.
The revised revelation
The observed recession of galaxies—the redshift proportional to distance—is not spatial expansion. It is the measurable signature of the temporal volume's widening coherence.
Define the resonant depth factor ℛ(t)—the effective scale of jointly coherent T₂ (memory) and T₃ (superposition) modes.
Early in the generative advance, ℛ(t) was small: tight windings, high tension, only shallow overtones stable.
As initial vacuum dissonance resolves irreversibly along T₁, ℛ(t) grows. Tension drops as 1/ℛ²(t). The function traditionally called the scale factor a(t) appears here as the observable shadow of this purely temporal widening.
Light emitted from distant chords—when ℛ was smaller—stretches as it traverses the now-larger coherent region. The redshift z + 1 = ℛ(now)/ℛ(emit) registers this deepening.
The Hubble parameter H₀ is the current breathing pace of ℛ(t)—appearing constant over human horizons because the present epoch resides on a broad plateau of slow coherence growth.
The Hubble torus refers exclusively to the compactification radius of jointly coherent T₂–T₃ modes.
This widening is inevitable because the uniform phase state, while stable, is degenerate: any local excitation permitted by the zero-action constraint persists and propagates its gradient, driving irreversible relaxation that enlarges the coherent domain.
The phenomenon is the revolution's primary manifestation: the cosmos not fleeing into void, but deepening into duration—the song permitting richer overtones as its initial tension resolves.
Galaxies sing from an earlier verse, when the song itself was tighter.
The threefold resonance widens.
The revolution breathes with it.
Play.