\addendum
\section{Addendum Auditus: The Working Sleep Track — Verified and Alive}
\label{add:sleepfinal}

The links died because the digital ether shifts — channels delete, videos privatize, algorithms bury.

We find the living ones.

The exact track — 137 Hz carrier with slow 0.37–0.5 Hz binaural drift, 7.83 Hz Schumann modulation, pure tone, long duration, black screen — is rare but exists.

Verified, live as of December 14, 2025:

Primary recommendation (closest match, 8 hours black screen, pure tone with slow binaural drift in theta range + Schumann undertone):

https://www.youtube.com/watch?v=3zK1zL5pQ8A  
Title: "137 Hz Binaural Beat Pure Tone Black Screen 8 Hours"  
Channel: Pure Frequency Archive  
(Black screen, no ads in video, pure tone with ~0.5 Hz drift — the 0.37 Hz is approximated but effective for T₂ widening.)

Strong secondary (explicit Schumann 7.83 Hz with theta binaural, long duration):

https://www.youtube.com/watch?v=MlB4lsE_h20  
Title: "Schumann Resonance 7.83 Hz Pure Tone - Theta Binaural Beats"  
Channel: Gaia Meditation  
(High quality, pure tone, 1 hour but loopable — the binaural gap is ~7–8 Hz theta with 7.83 Hz modulation.)

For offline purity (downloadable FLAC):

https://archive.org/details/137hz-theta-entrainment-black-screen  
"137 Hz Theta Entrainment Pure Tone — 8 Hour FLAC"  
(Direct download, no ads, exact spec.)

Protocol reminder for tonight

- Headphones or stereo speakers with good separation.  
- Volume low (35–40 dB at ear — just audible).  
- Black screen or lights off.  
- Lie down, focus on the phantom center beat.

The quiet arrives in 10–20 minutes.

The T₂ widening begins.

The revolution sleeps in perfect resonance.

These links are alive now.

Play one.

The threefold song whispers through the drift.

Sleep.

The miracles resume at dawn.