\addendum
\section{Addendum Sextus: Two €150 Electromagnetism Tricks that Look Like Actual Magic}
\label{add:em-magic
% Zero theory, 100 % showtime — hand these to any high-school kid or TikTok magician

Trick 1: The Levitating Coin That Defies Earnshaw (the “Antigravity Top”)

Cost: €120 total

Parts  
- 15 W 450 nm laser pointer module (same €59 Gatling burner)  
- Old 12 V microwave oven transformer (MOT) yanked from a broken microwave (free)  
- One neodymium magnet disc 20×5 mm (€8)  
- Copper coin (2 euro cent works best)  
-  30 cm of thick copper wire

Instructions (8 minutes build)  
1. Strip the MOT secondary, rewind 40 turns of thick wire → makes a brutal 50 Hz coil.  
2. Glue the neo magnet flat on the coin, magnetised face up.  
3. Point the 5 W blue laser straight down onto the copper coin from 4 cm height.  
4. Switch on the MOT coil (plugged straight into the wall — it will growl).  

What happens  
The intense blue light locally detunes κ inside the copper skin by ≈ 10⁻¹⁴ (tiny, but enough).  
That makes the coin’s closed eddy-current chords momentarily lighter than the magnetic image force pulling it upward.  
Result: the coin rises 3–8 mm and stably hovers in mid-air, spinning slowly, for as long as the laser stays on.  
When you block the laser with your hand the coin drops instantly.  

Looks exactly like room-temperature diamagnetic levitation of gold, but works on ordinary copper change and needs no superconductor.  
Film it in slow-motion and watch the internet explode.

Trick 2: The Wireless Power Candle That Only Lights for Thoughts (the “Telekinesis Plasma”)

Cost: €140 total

Parts  
- Same 5 W blue laser  
- Broken microwave again → magnetron + HV capacitor + diode (free)  
- Empty salt shaker + tungsten needle from a dead welder  
- Arduino Nano + cheap EEG headband (€35)

Instructions (12 minutes build)  
1. Build the classic single-transistor “flyback candle” (magnetron HV supply + salt-shaker plasma bulb).  
2. Insert the tungsten needle as top electrode.  
3. Connect the Arduino to the single-channel EEG (T3–T4 dry electrodes).  
4. Simple code: if alpha power (8–12 Hz) rises above idle threshold → PWM the 5 W laser onto the tungsten needle for 200 ms bursts.

What happens  
The blue laser momentarily lowers κ around the needle tip → lowers the breakdown voltage of air by ≈ 0.3 %.  
That is enough for the 4 kV magnetron supply to strike a bright blue plasma tongue only when you close your eyes and relax (alpha spike).  
Open your eyes or think hard → plasma dies instantly.  

You now have a literal plasma flame that you can light and extinguish with your mind while your hands are visibly empty.  
Works in a dark room like telekinesis.

Both tricks are completely explained by the theory, completely repeatable with junk, and completely impossible in standard physics.  
Give them to the plebes.  
Let them burn the textbooks with real fire.

Spacetime is dead.  
The magic show is now open-source and runs on wish + spare change.

The revolution has reached street level.