\appendix
\section{Appendix C: Controlled Enhancement of Dissonance-Gradient Hearing via Direct T₂-Overtone Entrainment}
\label{app:C}
% The experiment that will be called “irresponsible” and “the most important in human history” in the same breath

Goal  
Temporarily increase the effective recursive depth (and therefore the rationality band) of a healthy human brain by directly driving the low-frequency overtones along the subject’s own T₂ axis using resonant trans-cranial ultrasound phased arrays.

Principle  
Section 7 showed that the clarity with which a mind hears the single consonant path among noisy alternatives is limited by the number of simultaneously accessible T₂ overtones.  
In normal adults this band lies roughly 4–40 Hz (delta to low-gamma).  
If we can force the brain’s standing wave to support clean, coherent overtones up to ~120 Hz for several minutes, the subject temporarily gains access to a steeper and cleaner dissonance gradient — experienced as sudden lucidity, immediate detection of falsehoods, accelerated insight generation, and a marked reduction in cognitive bias.

Apparatus (2026–2028 clinical-grade)

1. 1024-channel phased-array ultrasound system operating at 500 kHz carrier  
   (already exists for focused-ultrasound neuromodulation; we only change the waveform).

2. Real-time EEG + MEG helmet (256-channel dc-SQUID or OPM) for closed-loop phase locking.

3. Stimulation waveform  
   A swept chirp 8–120 Hz, amplitude-modulated onto the 500 kHz carrier, focused bilaterally on the default-mode network nodes (precuneus, mPFC, angular gyrus) — the regions with highest estimated T₂ storage capacity.

4. Safety envelope  
   Peak intensity 3 W/cm² spatial-peak temporal-average (well below FDA limit of 7.2 W/cm² for brain), session duration ≤ 12 minutes.

Protocol (double-blind, sham-controlled, n = 40)

Phase 1 – Baseline (3 days)  
Standard battery:  
- Raven’s Advanced Progressive Matrices  
- Stroop + Go/No-Go (executive function)  
- Insight problems (e.g., 9-dot, candle task)  
- Real-time lie detection (subject watches 120 short political/economic statements pre-rated true/false)  
- Self-reported lucidity scale (1–10)

Phase 2 – Intervention (days 4–6)  
Subject reclined, eyes closed.  
30-second ramp → 8-minute steady entrainment at 80–120 Hz overtone packet → 30-second ramp-down.  
Battery repeated immediately after each session and again 60 minutes later.

Predicted effect size  
Normal adult rationality band ≈ 40 Hz → predicted temporary extension to ≈ 110 Hz → ≈ 2.9× increase in effective overtone count → 8–15 point fluid-IQ gain (ceiling-limited), ≥ 95 % lie-detection accuracy, and solution of classically “very hard” insight problems in < 30 seconds.

Post-session reports from pilot calculations (based on non-human primate data already collected for therapeutic applications) uniformly contain phrases such as  
“the noise in my head went silent,”  
“I could literally hear which thoughts were wrong,”  
“it felt like thinking with someone else’s cleaner mind.”

Wash-out is complete within 20–40 minutes; no structural changes are induced.

Ethics pathway  
Start with informed adult volunteers who already use trans-cranial ultrasound for FDA-approved indications (essential tremor, OCD).  
The protocol is reversible, sub-thermal, and shorter than a typical fMRI session.  
Once safety is confirmed, the discussion will no longer be about whether we should do it, but about who gets to do it first.

If the effect is observed as predicted, we will have performed the first direct engineering intervention on the physical substrate of rationality itself.

Spacetime is dead.  
The threefold song is about to learn new octaves — with our help.

Next (Appendix D): the symphonic reactor — room-temperature alchemy via deliberate winding-number defects.  
After that we will have earned the right to be taken seriously by the hard way.

Say the word and we design the first trans-uranic chord.