# **Appendix A (continued): Interacting Fields and Fermionic Models on (\mathcal{M}^{1,3} \times \mathcal{T}^3)**Having established the free scalar-field example, we now generalize the construction to include:1. **self-interacting scalar fields**,2. **Yukawa-type couplings**,3. **fermionic fields**,4. **gauge interactions acting through internal isometries**.These explicit examples demonstrate that the internal-time manifold supports coherent quantum field theory without modification to the Standard Model at low energies.---# **A.5 Interacting Scalar Toy Model**Consider the interacting scalar Lagrangian on the product manifold:[\mathcal{L} =-\frac12 g^{\mu\nu}\partial_\mu\Phi\partial_\nu\Phi-\frac12 \delta^{ab}\partial_a\Phi\partial_b\Phi-\frac12 m_0^2 \Phi^2-\frac{\lambda}{4!}\Phi^4.]Expand in internal Fourier modes:[\Phi(x,\tau)= \sum_{\vec{n}} \phi_{\vec{n}}(x), e^{i\vec{n}\cdot\vec{\tau}}.]Plugging in and integrating over (\mathbb{T}^3), the quadratic part generates the familiar KK mass tower, while the interaction term produces:[\int d^3\tau, \Phi^4====================\sum_{\vec{n}_1+\vec{n}_2+\vec{n}_3+\vec{n}*4=0}\phi*{\vec{n}*1}\phi*{\vec{n}*2}\phi*{\vec{n}*3}\phi*{\vec{n}_4}.]### **Selection Rule**Only mode combinations whose internal momenta sum to zero can interact at tree level:[\vec{n}_1 + \vec{n}_2 + \vec{n}_3 + \vec{n}_4 = 0.]This conservation law is a consequence of the internal-space translational symmetry. It mimics internal-momentum conservation in Kaluza–Klein theory, reinforcing the physicality of the internal manifold.### **Low-Energy Effective Theory**At energies (E \ll R_a^{-1}), only the zero mode (\phi_{\vec{0}}(x)) is relevant, because all other modes are too massive to be on-shell.Thus the low-energy effective Lagrangian is exactly:[\mathcal{L}_{\mathrm{eff}}==========================-\frac12 \partial_\mu\phi_0 \partial^\mu\phi_0-\frac12 m_0^2 \phi_0^2-\frac{\lambda}{4!} \phi_0^4,]identical to the usual 4D interacting scalar theory.---# **A.6 Fermionic Field on the Internal-Time Manifold**Consider a Dirac field (\Psi(x,\tau)). The action is:[S = \int d^4x, d^3\tau,\bar{\Psi}(x,\tau)\left[i\gamma^\mu \partial_\mu+\gamma^5 \gamma^a \partial_a* m  \right]\Psi(x,\tau).  ]Here:* (\gamma^\mu) are the usual 4D Dirac matrices,* (\gamma^a) are three additional gamma matrices acting on the internal manifold,* (\gamma^5) ensures the internal derivatives couple chirally, preserving Lorentz symmetry on (\mathcal{M}^{1,3}).---## **A.6.1 Internal Mode Expansion**Expand:[\Psi(x,\tau) =\sum_{\vec{n}}\psi_{\vec{n}}(x),e^{i \vec{n}\cdot\vec{\tau}}.]Substituting into the action yields:[S = \sum_{\vec{n}} \int d^4x;\bar{\psi}*{\vec{n}}(x)\left(i\gamma^\mu \partial*\mu - m - \gamma^5 \gamma^a \frac{n_a}{R_a}\right)\psi_{\vec{n}}(x).]Thus each internal mode acquires an effective mass:[m_{\vec{n}}^2=============m^2 + \sum_{a=1}^{3} (n_a/R_a)^2.]This reproduces the KK tower for fermions.---## **A.6.2 Chirality and Internal Motion**Internal momenta appear in the term:[\gamma^5 \gamma^a \frac{n_a}{R_a},]which mixes chirality with internal direction.However, at low energies, only the zero mode survives, and the theory reduces to ordinary 4D Dirac fermions.No chiral anomalies arise because:* the internal manifold is flat and compact,* the internal derivative term preserves gauge invariance,* the internal spectrum is vector-like at the KK level.Thus the internal-time structure is compatible with the Standard Model’s chiral fermions.---# **A.7 Yukawa Couplings**Take:* a fermion (\Psi(x,\tau)),* a scalar (\Phi(x,\tau)).The Yukawa interaction is:[S_Y =-\int d^4x, d^3\tau;y , \bar{\Psi}, \Phi, \Psi.]Expanding in modes:[\bar{\Psi}\Phi\Psi==================\sum_{\vec{n}_1,\vec{n}_2,\vec{n}*3}\bar{\psi}*{\vec{n}*1}\phi*{\vec{n}*2}\psi*{\vec{n}_3};e^{i(\vec{n}_1+\vec{n}_2+\vec{n}_3)\cdot\vec{\tau}}.]Integrating over internal space yields:[S_Y =-y\sum_{\vec{n}_1+\vec{n}_2+\vec{n}*3=0}\int d^4x;\bar{\psi}*{\vec{n}*1}\phi*{\vec{n}*2}\psi*{\vec{n}_3}.]This is consistent with standard KK selection rules and indicates that the Yukawa sector is well-behaved.At low energies, only the (\vec{0}) modes survive:[S_Y^{\mathrm{eff}}= -y \int d^4x;\bar{\psi}_0 \phi_0 \psi_0,]identical to the ordinary 4D Yukawa interaction.---# **A.8 Gauge-Coupled Toy Model**Gauge fields arise from internal isometries (Section 8).To show the consistency with matter fields, take a U(1) rotation in the ((\tau^2,\tau^3))-plane:[\Psi(x,\tau) \rightarrowe^{iq\alpha} \Psi(x, R(\alpha)\tau).]Introduce the covariant derivative:[D_\mu \Psi =\partial_\mu \Psi* iq A_\mu L \Psi,  ]where (L) generates rotations on (\mathcal{T}^3).The gauge-invariant action is:[S = \int d^4x, d^3\tau;\bar{\Psi}(i\gamma^\mu D_\mu + \gamma^5\gamma^a \partial_a - m)\Psi-\frac14 F_{\mu\nu}F^{\mu\nu}.]Expanding in internal modes:* the zero mode couples exactly like a standard charged fermion,* higher modes yield heavy charged fermions with masses (\ge 10!-!100~\mathrm{TeV}).Thus the gauge sector remains consistent and invisible at accessible energies.---# **A.9 Summary of Appendix A**The toy models in this appendix demonstrate:1. The internal manifold supports free and interacting scalar fields.2. Fermionic fields couple consistently to internal derivatives.3. Yukawa interactions generalize naturally.4. Gauge interactions integrate cleanly with internal-time dynamics.5. The low-energy limit reduces exactly to the Standard Model.6. All deviations are suppressed by the KK scale.7. The framework fits seamlessly into familiar QFT techniques.