# **Appendix C. Relation to the Page–Wootters Relational-Time Mechanism**The Page–Wootters (PW) mechanism provides a relational account of time in which the universe as a whole is in a stationary (timeless) state, while subsystems exhibit effective time evolution through correlations with a designated “clock” subsystem. In this appendix we show how the tri-axial internal-time construction fits into and generalizes the PW framework.---## **C.1 The Standard Page–Wootters Setup**In the PW formalism, one partitions the total Hilbert space into a clock sector (\mathcal{H}_C) and a system sector (\mathcal{H}_S):[\mathcal{H}_{\mathrm{tot}} = \mathcal{H}_C \otimes \mathcal{H}_S.]A total Hamiltonian is assumed of the form:[\hat{H}_{\mathrm{tot}} = \hat{H}_C \otimes \mathbb{I}_S + \mathbb{I}_C \otimes \hat{H}_S,]together with the constraint:[\hat{H}_{\mathrm{tot}} |\Psi\rangle = 0,]which is a Wheeler–DeWitt-type equation. The state (|\Psi\rangle \in \mathcal{H}_{\mathrm{tot}}) is stationary with respect to any external time.One then chooses a basis of “clock states” ({|t\rangle}) in (\mathcal{H}_C) such that:[\hat{H}_C |t\rangle = i \frac{\partial}{\partial t} |t\rangle,]and defines conditional system states:[|\psi(t)\rangle_S = \frac{\langle t| \Psi \rangle}{\sqrt{P(t)}},]where[P(t) = \langle \Psi |t\rangle\langle t | \Psi \rangle]is the probability density for the clock reading (t).Under appropriate assumptions (e.g. clock behaving semiclassically), one shows:[i \frac{\partial}{\partial t} |\psi(t)\rangle_S = \hat{H}_S |\psi(t)\rangle_S,]i.e., the usual Schrödinger equation emerges as a conditional statement about correlations between the clock and the system.---## **C.2 Internal Time as a Three-Dimensional Clock Hilbert Space**In the tri-axial framework, the internal temporal manifold (\mathcal{T}^3) provides a natural candidate for the clock sector. For a given physical system, we identify:[\mathcal{H}_C = L^2(\mathcal{T}^3), \qquad\mathcal{H}_S = L^2(\mathcal{M}^{1,3}) \ \text{(or an appropriate subsystem thereof)}.]The total Hilbert space is thus:[\mathcal{H}_{\mathrm{tot}}= L^2(\mathcal{T}^3) \otimes \mathcal{H}_S,]which matches the structure used in the main text:[\mathcal{H} = L^2(\mathcal{M}^{1,3}) \otimes L^2(\mathcal{T}^3),]up to the choice of which sector is regarded as “clock” and which as “system”.We choose internal coordinates (\tau^a) and basis states (|\tau\rangle = |\tau^1, \tau^2,\tau^3\rangle) satisfying:[\langle \tau | \tau' \rangle = \delta^{(3)}(\tau - \tau'),\qquad\int_{\mathcal{T}^3} d^3\tau, |\tau\rangle\langle \tau| = \mathbb{I}_C.]---## **C.3 The Hamiltonian Constraint in the Tri-Axial Model**From the main text, we take the total Hamiltonian:[\hat{H}_{\mathrm{tot}} = \hat{H}_C + \hat{H}_S,]with:[\hat{H}_C= -\frac{1}{2}\delta^{ab} \partial_a \partial_b + V_C(\tau),]acting on (\mathcal{H}_C), and (\hat{H}_S) acting on (\mathcal{H}_S) as the usual system Hamiltonian (or a Dirac/Klein–Gordon operator for relativistic fields). The physical state satisfies:[\hat{H}_{\mathrm{tot}} |\Psi\rangle = 0.]In the (|\tau\rangle) representation:[\Psi(\tau) := \langle \tau | \Psi \rangle \in \mathcal{H}_S,]and the constraint reads:[\left[-\frac{1}{2}\delta^{ab}\partial_a\partial_b* V_C(\tau)* \hat{H}_S  \right] \Psi(\tau) = 0.  ]This is a functional-differential equation for (\Psi(\tau)), with (\hat{H}_S) acting on the system degrees of freedom and (\partial_a) acting on the internal coordinates.---## **C.4 Emergent 1D Time from a 3D Internal Manifold**To recover a single emergent time parameter (t), one typically requires that the internal-clock wavefunction is strongly peaked along a trajectory in (\mathcal{T}^3). For example, suppose that:* The clock potential (V_C(\tau)) is such that the clock wavepacket follows a semiclassical path (\tau^3 \approx \omega t), with (\tau^1,\tau^2) remaining narrowly distributed around fixed values.* The solution (\Psi(\tau)) can be approximated as:[\Psi(\tau) \approx \chi(\tau^3) \otimes |\psi(\tau^3)\rangle_S,]with (\chi) sharply peaked and slowly dispersing.Then, integrating over (\tau^1,\tau^2) and projecting onto the dominant clock trajectory, one effectively reduces the PW construction to a one-dimensional internal clock coordinate (\tau^3), with an emergent parameter (t \sim \tau^3/\omega).Formally, define:[|\psi(t)\rangle_S=================\frac{\int d\tau^1 d\tau^2, \langle \tau^1,\tau^2,\tau^3(t)|\Psi\rangle}{\sqrt{P(t)}},]with:[P(t) = \int d\tau^1 d\tau^2,|\langle \tau^1,\tau^2,\tau^3(t)|\Psi\rangle|^2.]Under the assumption that (V_C(\tau)) yields a semiclassical evolution along (\tau^3) and that the dependence on (\tau^1,\tau^2) is weak (or can be averaged out), one obtains:[i \frac{\partial}{\partial t} |\psi(t)\rangle_S = \hat{H}_S |\psi(t)\rangle_S,]i.e., the standard Schrödinger equation.Thus, the traditional PW construction reappears as a **special case** of the tri-axial model, corresponding to a clock state localized along a one-dimensional trajectory within (\mathcal{T}^3).---## **C.5 Multi-Parameter Conditional Evolutions**Because the internal space has three dimensions, the tri-axial model allows more general conditional structures than the original PW scheme. Instead of selecting a single internal coordinate as “time,” one may consider conditional evolutions along surfaces or curves more complex than (\tau^3 = \omega t).For instance, define a conditional state:[|\psi(\sigma)\rangle_S======================\frac{\int d^3\tau\ f(\tau;\sigma), \langle \tau | \Psi \rangle}{\sqrt{P(\sigma)}},]where (f(\tau;\sigma)) is a weight function selecting a family of internal configurations parametrized by (\sigma). Different choices of (f) correspond to different “gauges” for internal time:* Lines primarily along (\tau^1) correspond to histories emphasizing retentive structure.* Lines along (\tau^2) emphasize superpositional/counterfactual structure.* Lines along (\tau^3) emphasize anticipatory or projective structure.In all cases, the constraint equation ensures that these conditional states obey effective Schrödinger-like equations with different emergent Hamiltonians, reflecting the chosen slicing of internal time.The standard PW time parameter is thus a particular, physically convenient choice of foliation in the internal temporal manifold.---## **C.6 Decoherence and Preferred Internal Time Directions**In practice, not all slicings of (\mathcal{T}^3) yield stable or experimentally accessible “times.” Interactions with the environment, internal dynamics, and stability conditions can select preferred directions in internal-time space, analogous to how decoherence selects preferred pointer bases in ordinary Hilbert space.The tri-axial structure allows:* **preferred axes** in (\mathcal{T}^3) to emerge from dynamical and environmental factors,* **robust emergent time parameters** associated with those axes,* **context-dependent internal time choices** within a single global framework.In this sense, the tri-axial model extends the PW mechanism from a single abstract clock to a structured internal temporal geometry in which practical time parameters emerge as stable directions.---## **C.7 Summary of the Relation to Page–Wootters**1. The tri-axial internal-time construction realizes the PW idea with a **three-dimensional internal clock space** instead of a single clock variable.2. The total state obeys a **timeless Hamiltonian constraint**, as in PW and canonical quantum gravity.3. **Single-time Schrödinger dynamics** arise by conditioning on suitable trajectories or foliations in (\mathcal{T}^3).4. The original PW formalism is recovered as a special case where the internal state is sharply localized along a one-dimensional path.5. The additional internal dimensions allow **richer relational structures** and potential new insights into decoherence, emergent classicality, and the selection of effective time parameters.This completes the technical bridge between the tri-axial internal-time framework and existing relational-time approaches.