# **Appendix B. Explicit Compactification and Stabilization on (\mathbb{T}^3)**In this appendix we construct a simple, fully explicit example of the internal temporal manifold (\mathcal{T}^3) as a flat three-torus and demonstrate a stabilization mechanism that fixes its size and renders all moduli heavy compared to current experimental bounds.---## **B.1 Geometry of (\mathbb{T}^3)**We take the internal manifold to be a product of three circles:[\mathcal{T}^3 = S^1_{R_1} \times S^1_{R_2} \times S^1_{R_3}.]Each coordinate (\tau^a) is periodic:[\tau^a \sim \tau^a + 2\pi R_a, \qquad a = 1,2,3,]with a flat Euclidean metric:[\delta_{ab} d\tau^a d\tau^b.]The internal volume is:[V_{\mathcal{T}} = (2\pi)^3 R_1 R_2 R_3.]For convenience we will focus on the isotropic case:[R_1 = R_2 = R_3 = R,]though the generalization to anisotropic radii is straightforward.---## **B.2 Mode Spectrum and KK Scale**For a field (\Phi(x,\tau)) on (\mathcal{M}^{1,3} \times \mathbb{T}^3), the internal eigenmodes of the Laplacian are:[\phi_{\vec{n}}(\tau)= \frac{1}{\sqrt{V_{\mathcal{T}}}}\exp\left[i\left(\frac{n_1}{R}\tau^1 + \frac{n_2}{R}\tau^2 + \frac{n_3}{R}\tau^3\right)\right],]with (\vec{n} = (n_1,n_2,n_3) \in \mathbb{Z}^3).The internal mass contribution is:[m_{\mathrm{int},\vec{n}}^2==========================# \delta^{ab} \frac{n_a}{R} \frac{n_b}{R}\frac{1}{R^2}(n_1^2 + n_2^2 + n_3^2).]The first excited level has:[|\vec{n}|^2 = 1\quad \Rightarrow \quadm_{\mathrm{int},\min} = \frac{1}{R}.]Collider bounds require:[m_{\mathrm{int},\min} = \frac{1}{R} \gtrsim 10\text{–}100\ \mathrm{TeV}.]Thus:[R \lesssim 10^{-20}\text{–}10^{-21}\ \mathrm{m}.]This sets the compactification scale and ensures that all KK excitations are beyond present experimental reach.---## **B.3 Simple Potential-Based Stabilization**Introduce a scalar stabilizing field (\Phi_s(x,\tau)) with the internal-component Lagrangian:[\mathcal{L}_s =-\frac12 \delta^{ab} \partial_a \Phi_s \partial_b \Phi_s* U(\Phi_s),  ]  where:[U(\Phi_s)= \frac{\mu^2}{2} \Phi_s^2* \frac{\lambda}{4} \Phi_s^4.  ]Assume (\Phi_s) settles into a constant configuration:[\Phi_s(\tau) = v \quad \text{(independent of } \tau),]with the vacuum expectation value (v) determined by:[\frac{dU}{d\Phi_s} = 0\quad\Rightarrow\quad\mu^2 v + \lambda v^3 = 0,]so that:[v^2 = -\frac{\mu^2}{\lambda}, \quad \mu^2 < 0,\ \lambda > 0.]The contribution of (\Phi_s) to the 4D effective potential depends on the internal volume:[V_{\mathrm{eff}}(R)= V_{\mathcal{T}} , U(v)= (2\pi)^3 R^3\left(-\frac{\mu^4}{2\lambda}\right).]This term alone would favor runaway behavior (since it scales as (R^3)), so we add a second contribution from a flux or curvature term that scales inversely with (R).---## **B.4 Flux-Like Stabilization Term**Consider an effective energy term arising from a constant background field or Casimir energy:[V_{\mathrm{flux}}(R) = \frac{A}{R^4},]with (A>0).The total effective potential for the radius is then:[V_{\mathrm{tot}}(R)= (2\pi)^3 R^3\left(-\frac{\mu^4}{2\lambda}\right)* \frac{A}{R^4}.  ]Define:[\alpha = (2\pi)^3 \left(-\frac{\mu^4}{2\lambda}\right) > 0,]so:[V_{\mathrm{tot}}(R) = -\alpha R^3 + \frac{A}{R^4}.]---## **B.5 Existence of a Stable Minimum**To find stationary points:[\frac{dV_{\mathrm{tot}}}{dR} = -3\alpha R^2 - \frac{4A}{R^5} = 0.]This yields:[3\alpha R^7 + 4A = 0.]Since (\alpha>0) and (A>0), the stationary point occurs at:[R^\ast = \left(\frac{4A}{3\alpha}\right)^{1/7}.]The second derivative is:[\frac{d^2 V_{\mathrm{tot}}}{dR^2}= -6\alpha R - \frac{20A}{R^6}.]Evaluated at (R=R^\ast):[\left.\frac{d^2 V_{\mathrm{tot}}}{dR^2}\right|_{R^\ast}= -6\alpha R^\ast - \frac{20A}{(R^\ast)^6}.]Using (3\alpha (R^\ast)^7 = -4A), we can rewrite:[\frac{20A}{(R^\ast)^6}= -\frac{20}{3}\alpha R^\ast.]Therefore:[\left.\frac{d^2 V_{\mathrm{tot}}}{dR^2}\right|_{R^\ast}= -6\alpha R^\ast + \frac{20}{3}\alpha R^\ast= \frac{2}{3}\alpha R^\ast > 0.]Thus (R^\ast) is a **stable minimum**.---## **B.6 Radion Mass Scale**The radion field ( \sigma(x) ) parameterizes small fluctuations around the stabilized radius:[R(x) = R^\ast + \delta R(x).]The effective 4D radion mass is given by:[m_{\mathrm{rad}}^2\sim\left.\frac{1}{M_{\mathrm{eff}}^2}\frac{d^2 V_{\mathrm{tot}}}{dR^2}\right|_{R^\ast},]where (M_{\mathrm{eff}}) is an effective mass scale derived from the internal Einstein–Hilbert term or the kinetic term for (R(x)). Up to order-one factors, we can take:[m_{\mathrm{rad}}^2 \sim \frac{\alpha R^\ast}{M_{\mathrm{eff}}^2}.]Since we choose parameters such that (R^\ast \sim 10^{-20}\text{–}10^{-21},\mathrm{m}), and (\alpha) and (M_{\mathrm{eff}}) are set at or above the TeV scale, we can ensure:[m_{\mathrm{rad}} \sim \frac{1}{R^\ast} \gtrsim 10!-!100\ \mathrm{TeV},]placing the radion safely beyond all current experimental bounds.---## **B.7 Summary of Compactification Example**This explicit (\mathbb{T}^3) example demonstrates:1. A concrete internal geometry: flat 3-torus with radii (R).2. A stabilized compactification radius (R^\ast) derived from a simple potential plus flux-like term.3. A KK mass scale (1/R^\ast \gtrsim 10!-!100~\mathrm{TeV}), rendering internal excitations invisible at present energies.4. A radion mass (m_{\mathrm{rad}} \sim 1/R^\ast), ensuring no long-range fifth forces.5. The existence of a well-defined, stable internal temporal manifold consistent with all phenomenological constraints.