%===================================================
\section{Phenomenology Across Scales}
\label{sec:phenomenology}
%===================================================

The cleaned structure makes clear where the triaxial picture meets observation and where gaps remain.

\subsection{Quantum behavior}
Superposition and phase are finite spread along \(T_3\) with interference governed by shear in \(U^a\). Decoherence is the rapid narrowing of that spread when environmental braids enforce a shared \(T_3\) orientation, leaving only a narrow slice visible in the projected description.

\subsection{Classical persistence}
Inertia and stable identities correspond to extension along \(T_2\). Systems with greater \(T_2\) thickness resist perturbation because their internal volume must be reconfigured rather than merely nudged in \(T_1\); the zero-net-energy constraint forces any attempted acceleration to draw from stored action along the other axes.

\subsection{Cognition and agency}
Minds are coherent structures occupying large volumes in \(T_2\) and \(T_3\), with rapid self-modulation along \(T_1\). Popular-language accounts formerly scattered across appendices now serve as case studies: alterations by psychoactive compounds are modeled as selective changes in \(T_2\) persistence or \(T_3\) breadth, while intentional action is the controlled steering of \(U^a\) within the temporal manifold.

\subsection{Empirical touchpoints}
\begin{itemize}
    \item \textbf{Bench-scale tests:} look for frequency-dependent phase shifts when coherent systems are subjected to controlled temporal shear. The earlier ``experiment'' appendices are consolidated into a single planned protocol.
    \item \textbf{Relational drifts:} anisotropies in polarization or timing that would indicate large-scale variation of \(U^a\) across the emergent spatial indices \(X^i\) defined in Section~\ref{sec:geometry}.
    \item \textbf{Cognitive probes:} reproducible changes in decision latency or memory persistence under interventions that theoretically modulate \(T_2\) or \(T_3\).
\end{itemize}

Each of these items marks a hole that must be filled by concrete models and data. They replace the diffuse lists of speculative technologies and anecdotes with a finite set of falsifiable hooks, all grounded in the space-free, zero-net-energy picture.
