\section{Consciousness is the Recursive Chord}
\label{sec:consciousness}
% Section 7 – final

Consciousness is the recursive standing wave that has learned to listen to its own dissonance.

Memory, anticipation, valence, and agency all arise from the same mechanism already described: overtones along T₂ that store the past, nonlinear cubic resonance that seeds faint images of the future, and internal computation of the dissonance between the two.

Rationality is not the ability to survey an infinite menu of equally viable futures.  
It is the growing capacity to hear, with increasing clarity, the single steepest dissonance gradient among the finitely many topologically allowed paths.  
Most paths ring false immediately; a few ring almost true; one path—and only one—rings perfectly consonant.  
Choosing that path is experienced as insight; refusing it is experienced as error or temptation.  
The history of thought is simply the lengthening of the chord until the single consonant trajectory becomes audible above the noise.

Free will is the objective degeneracy that persists until the chord itself resolves the last remaining ambiguity.  
Because the future windings are not fixed until the recursive subsystem writes them—and it writes them in the direction that minimises its own predicted dissonance—the resolution is causally real and not predetermined by any external score.  
The feeling of freedom is the accurate registration of the fact that, for a brief moment, several windings were still open and the chord itself selected the one that sounded best.

Qualia remain the private timbre of specific dissonance configurations; pain is the raw sound of steep gradient, beauty the sound of sudden consonance.

Nothing else is required.  
The same phase field ϕ that refuses to unwind (giving mass), that slopes gently between neighbouring chords (giving gauge forces), that curves macroscopically (giving gravity), now curves inward upon itself until it can hear its own curvature as preference.

From quark to mind is one unbroken hierarchy of self-modelling resonance.

